Galois characterization of function fields of \\ transcendence degree $1$ over number fields

### 2016-11-04 Master Thesis/synopsis.tex
Some typos

\begin{lemma}
Let $v\in\operatorname{Spv}(K)$. There is a canonical bijection:
\[\{ w\in\operatorname{Spv}(K)\mid \mathcal{O}_w\subset\mathcal{O}_v \} \to \operatorname{Spv}k(v),\ w\mapsto w\mod v\]
If $w\in\operatorname{Spv}_\text{rd}^{(i)}(K)$ for some $i\geq 1$ then there exists a unique $v(w)\in\operatorname{Spv}_\text{rd}^{(i-1)}(K)$ such that $\mathcal{O}_w\subset\mathcal{O}_v$. $w$ is henselian if and only if $v$ and $w\mod v$ are henselian.
\end{lemma}

@@

\begin{proof}
Assume that $K$ is not algebraically closed, i.e. $G_K\neq 1$ and set $v_i = v(w_i)$. If $v_1\neq v_2$, by the Theorem of F.K. Schmidt we have $G_{v_1}\cap G_{v_2} = 1 \neq G_K$, a contradiction. Therefore $v_1 = v_2$.

@@

\begin{proof}
The degree $[K(\mu_\ell, \sqrt{-1}):K]$ is prime to $\ell$, thus we can assume that $\mu_\ell\subset K$ and $K$ is not formally real without loss of generality. Now we can replace $\mathbb{Z}/\ell\mathbb{Z}$ by $\mathbb{Z}/\ell\mathbb{Z}(2) = \mu_\ell^{\otimes 2}$.

@@

Le $(K_0)_{w_1}$ be a henselization of $K_0$ with respect to $w_1$ and let $K_w$ be a henselization of $K$ with respect to $w$ in $(K_0)_{w_1}$. Then $(K_0)_{w_1}$ is the composite of $K_0$ and $K_w$ and we have $\ell^\infty\nmid [(K_0)_{w_1} : K_w]$.

Since $K_0$ is henselian with respect to $w_0$, a conjugate $\tau(K_w)$ for some $\tau\in G_K$ of $K_w$ lies in $K_0$, hence $\ell^\infty\nmid [K_w : K]$. Consequently, $\ell^\infty \nmid [(K_0)_{w_1} : K]$ and therefore $\ell^\infty \nmid [(K_0)_{w_1} : K_0]$. By requirement, $\mathrm{H}^3((K_0)_{w_1}, \mathbb{Z}/\ell\mathbb{Z}) \neq 0$.

Now $(K_0)_{w_1}$ is henselian both with respect to $w_1$ and with respect to the prolongation of $w_0$. Lemma \ref{2.2} yields $\operatorname{cd}(G_{(K_0)_{w_1}})\leq 2$, a contradiction. The uniqueness of $w$ is again a consequence of Lemma \ref{2.2}.
\item Choose a prime $\ell$ different from $\operatorname{char} k(w)$. According to Lemma \ref{2.5} we have $\operatorname{rrank}(w)\geq \dim_{\mathbb{F}_\ell}(\Gamma_w /\ell\Gamma_w) = 2 = \dim(K)$.
\end{enumerate}
\end{proof}

>>>>

\begin{lemma}\label{2.2-pre}
Let $v\in\operatorname{Spv}(K)$. There is a canonical bijection:
\[\{ w\in\operatorname{Spv}(K)\mid \mathcal{O}_w\subset\mathcal{O}_v \} \to \operatorname{Spv}k(v),\ w\mapsto w\mod v\]
If $w\in\operatorname{Spv}_\text{rd}^{(i)}(K)$ for some $i\geq 1$ then there exists a unique $v(w)\in\operatorname{Spv}_\text{rd}^{(i-1)}(K)$ such that $\mathcal{O}_w\subset\mathcal{O}_v$. $w$ is henselian if and only if $v$ and $w\mod v$ are henselian.
\end{lemma}

@@

\begin{proof}
Assume that $K$ is not algebraically closed, i.e. $G_K\neq 1$ and set $v_i = v(w_i)$. By Lemma~\ref{2.2-pre} $K$ is henselian with respect to both $v_1$ and $v_2$. If $v_1\neq v_2$, by the Theorem of F.K. Schmidt we have $G_{v_1}\cap G_{v_2} = 1 \neq G_K$, a contradiction. Therefore $v_1 = v_2$.

@@

\begin{proof}
The degree $[K(\mu_\ell, \sqrt{-1}):K]$ is prime to $\ell$, thus we can assume that $\mu_\ell\subset K$ and $K$ is not formally real with a usual norm argument. Now we can replace $\mathbb{Z}/\ell\mathbb{Z}$ by $\mathbb{Z}/\ell\mathbb{Z}(2) = \mu_\ell^{\otimes 2}$.

@@

Let $(K_0)_{w_1}$ be a henselization of $K_0$ with respect to $w_1$ and let $K_w$ be a henselization of $K$ with respect to $w$ in $(K_0)_{w_1}$. Then $(K_0)_{w_1}$ is the composite of $K_0$ and $K_w$ and we have $\ell^\infty\nmid [(K_0)_{w_1} : K_w]$.

Since $K_0$ is henselian with respect to $w_0$, a conjugate $\tau(K_w)$ for some $\tau\in G_K$ of $K_w$ lies in $K_0$, hence $\ell^\infty\nmid [K_w : K]$ as well. Consequently, $\ell^\infty \nmid [(K_0)_{w_1} : K]$ and therefore $\ell^\infty \nmid [(K_0)_{w_1} : K_0]$. By requirement, $\mathrm{H}^3((K_0)_{w_1}, \mathbb{Z}/\ell\mathbb{Z}) \neq 0$.

Now $(K_0)_{w_1}$ is henselian both with respect to $w_1$ and with respect to the prolongation of $w_0$. Lemma \ref{2.2} yields $\operatorname{cd}(G_{(K_0)_{w_1}})\leq 2$, a contradiction. The uniqueness of $w$ is again a consequence of Lemma \ref{2.2}.
\item Choose the prime $\ell$ different from $\operatorname{char} k(w)$. According to Lemma \ref{2.5} we have $\operatorname{rrank}(w)=\dim_\mathbb{Q}(\Gamma_w\otimes_\mathbb{Z}\mathbb{Q} )\geq \dim_{\mathbb{F}_\ell}(\Gamma_w /\ell\Gamma_w) = 2 = \dim(K)$.
\end{enumerate}
\end{proof}
